Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-13 paper/comoving-followup.tex
conclusions - BOOM

\include{gitstuff}
\include{preamble}
\graphicspath{{figures/}}
\usepackage{microtype}

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
\newcommand{\npairsobs}{311}
\newcommand{\ncomoving}{127}

@@

\begin{abstract}
% Context
Widely-separated ($\gtrsim 0.1$~pc), coeval stars are powerful tools for
testing star-formation model predictions, improving stellar atmosphere models,
and inferring the mass distribution of the Milky Way.
They likely outnumber the Galactic open clusters by a large margin, but are
still expected to share a common origin and chemistry.
The kinematic properties of comoving pairs---including bound wide binaries---
serve as potentiometers for the smooth tidal field and mass perturbation
spectrum of the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood using
\project{Hipparcos}.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
false-positive rate at large separations is significant ($\sim 50\%$).
% Methods
Here, we present results from our own low-resolution radial-velocity survey of a
sample of the candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
pairs with separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information supports the belief that the two stars are
truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
\ncomoving of the \npairsobs\ comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution of velocity-differences as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

\subsection{Properties of genuine comoving pairs}\label{sec:genuine}

\figurename~\ref{fig:dx-dv} shows the magnitude of the 3D velocity difference
plotted against the 3D positional separation of the two stars in all observed
pairs.
We again plot the median of the velocity difference distribution, and here use
the Lutz-Kelker-corrected distance estimate (\citealt{Lutz:1973}) computed from
the stars' \tgas\ parallax measurements to compute the 3D positions of the two
stars.
Small, gray markers are pairs for which $\llrnew < \llrold$, i.e. where the
radial velocity measurements weaken or change our belief that a given pair is
comoving.
Larger, black markers are genuine comoving pairs ($\llrnew > \llrold$); note
that many large-separation pairs have been ruled-out as comoving, but
most small-separation pairs maintain their status as comoving.
This is also apparent in \figurename~\ref{fig:separation}, which shows
histograms of the spatial separation and sky separation for all pairs;
background, gray histogram shows all observed pairs, and foreground, black
histogram shows all pairs for which $\llrnew > \llrold$.

At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
genuinely comoving; interestingly, the separation distribution of the genuine
pairs appears to be uniform or possibly growing in number with increasing
separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
    the stars in all observed comoving star pairs (gray), and only those
    identified as genuinely comoving (black). Note that the contamination rate
    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
    significant number of genuine comoving pairs at these very wide separations.
    \emph{Right}: Distribution of chord lengths (tangential separations) for all
    comoving star pairs (gray) and only those identified as genuinely comoving
    (black).
    Noise in the parallax measurements makes the 3D separation distribution
    appear roughly uniform.
    Even the distribution in chord length appears to be approximately linear for
    large separations ($\gtrsim 0.1~\pc$).
    \label{fig:separation}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
    the density of all \tgas\ sources within 200 pc; similar to
    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
    pairs.
    Several interesting pairs that include main-sequence and giant component
    stars are highlighted in the text
    (\sectionname~\ref{sec:interesting-pairs} and
    \figurename~\ref{fig:highlighted-cmd}).
    \label{fig:genuine-cmd}}
\end{figure}

\subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}

@@

@@

HD 95088 HD 92278
Angular sep.: 4.49 deg
Chord length: 8.66 pc
3D sep.: 9.53 pc
parallaxes: 9.00 +/- 0.50, 9.32 +/- 0.51

This pair is separated by $8.7~\pc$ in chord length ($4.5^\circ$ on the sky) and
have consistent distances within their uncertainties of $\approx 110~\pc$.
From visual classification of the stars' location in the color--magnitude
diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
comprised of an A-type dwarf and a slightly less massive G-type subgiant.
If this is indeed the case, the two stars could therefore not be coeval, as the
more massive A star has not begun its post-main-sequence evolution.
One possible explanation is that this pair is a false-positive in three
dimensions, in which case they are unassociated but just appear to be comoving
with small relative velocity at present day.

Other possibilities stem from the possibility that the comoving pair is the
outcome of an interaction between more stars.
For example, an interaction between two wide binaries could result in an
exchange of members, a hardening of one resulting binary, and a widening of the
other binary (see, e.g., \citealt{Nathan??}).
Or, the initial system could have been an unstable three-body system that
resulted in a tight, inner binary and a wide outer binary;
if the inner binary were close enough, it could have merged and formed a blue
straggler star (e.g., \citealt{??}), leaving the wide binary untouched.

These explanations are all speculative at the moment; high-resolution
spectroscopy of these stars --- and precise measurements of the stellar
parameters and chemical abundances --- may help to distinguish these
possibilities.

@@

This pair is separated by $XX~\pc$ in chord length ($XX^\circ$ on the sky) and
have consistent distances within their uncertainties of $\approx XX~\pc$.
From visual classification of the stars' location in the color--magnitude
diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
comprised of a G-type subgiant and a G-type dwarf.

@@

TYC 3792-658-1 HD 54318
Angular sep.: 2.51 deg
Chord length: 8.45 pc
3D sep.: 8.73 pc
parallaxes: 5.20 +/- 0.23, 5.30 +/- 0.32

This pair is separated by $8.5~\pc$ in chord length ($2.5^\circ$ on the sky) and
have consistent distances within their uncertainties of $\approx 190~\pc$.

From visual classification of the stars' location in the color--magnitude
diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
comprised of a late-A-/early-B-type dwarf and a G-type dwarf.

\subsection{Combined with RAVE}\label{sec:rave}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
  \end{center}
  \caption{%
    \label{fig:separation-RAVE}}
\end{figure}

Caution about selection criteria, false-positives at large separation...

@@

Abundance of large-separation pairs

@@

@@

Low-resolution spectroscopic follow-up of comoving pairs.

Using relative RV's, confirm a sample of very wide separation comoving star pairs.

Either wide binaries, unbound binaries, or dissolving clusters

Need much larger sample with no cuts on separation or signal-to-noise to study true separation distribution

Gaia DR2 for G < 12, but really we need a concerted effort to obtain spectroscopy of wide pairs.

@@

Dave Zurek, Keith, Ruth, Kelle
John Brewer (Yale),
Jason Curtis (Columbia),
James Davenport (WWU) for PyDIS,
David W. Hogg (NYU),
Melissa Ness (MPIA)

This work is based on observations obtained at the MDM Observatory, operated by
Dartmouth College, Columbia University, Ohio State University, Ohio University,
and the University of Michigan.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.
This research has made use of the SIMBAD database, operated at CDS, Strasbourg,
France.
% The authors are pleased to acknowledge that the work reported on in this
% paper was substantially performed at the TIGRESS high performance computer
% center at Princeton University which is jointly supported by the Princeton
% Institute for Computational Science and Engineering and the Princeton
% University Office of Information Technology's Research Computing department.

>>>>

\include{gitstuff}
\include{preamble}
\graphicspath{{figures/}}
\usepackage{microtype}
\usepackage{multirow}

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\logg}{\ensuremath{\ensuremath{\log g}}}
\newcommand{\teff}{\ensuremath{\ensuremath{T_{\rm eff}}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}

\newcommand{\npairsobs}{311}
\newcommand{\ncomoving}{127}
\newcommand{\nraveoverlap}{154}
\newcommand{\nravecomoving}{59}
\newcommand{\ncomovingtotal}{186}

@@

\begin{abstract}
% Context
Widely-separated ($\gtrsim 0.1$~pc), coeval stars are powerful tools for
testing star-formation model predictions, improving stellar atmosphere models,
and inferring the mass distribution of the Milky Way.
They likely outnumber the Galactic open clusters by a large margin, but are
still expected to share a common origin and chemistry.
The kinematic properties of comoving pairs---including bound wide binaries---
serve as potentiometers for the smooth tidal field and mass perturbation
spectrum of the Milky Way.
Thus far, a few hundred confirmed comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood using
\project{Hipparcos}.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
false-positive rate at large separations is significant ($\sim 50\%$).
% Methods
Here, we present results from our own low-resolution radial-velocity survey of a
sample of the candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
pairs with separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information supports the belief that the two stars are
truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
radial velocity data as \llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
\ncomoving of the \npairsobs\ comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution of velocity-differences as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

\subsection{Confirmed comoving pairs}\label{sec:genuine}

\figurename~\ref{fig:dx-dv} shows the magnitude of the 3D velocity difference
plotted against the 3D positional separation of the two stars in all observed
pairs from this work.
We again plot the median of the velocity difference distribution, and here use
the Lutz-Kelker-corrected distance estimate (\citealt{Lutz:1973}) computed from
the stars' \tgas\ parallax measurements to compute the 3D positions of the two
stars.
Small, gray markers are pairs for which $\llrnew < \llrold$, i.e. where the
radial velocity measurements weaken or change our belief that a given pair is
comoving.
Larger, black markers are confirmed comoving pairs ($\llrnew > \llrold$); note
that many large-separation pairs have been ruled-out as comoving, but
most small-separation pairs maintain their status as comoving.
This is also apparent in \figurename~\ref{fig:separation}, which shows
histograms of the spatial separation and sky separation for all pairs;
background, gray histogram shows all observed pairs, and foreground, black
histogram shows all pairs for which $\llrnew > \llrold$.

At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
confirmed to be comoving; interestingly, the separation distribution of the
genuine pairs appears to be uniform or possibly growing in number with
increasing separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
    the stars in all observed comoving star pairs (gray), and only those
    confirmed as comoving (black).
    Note that the false-positive rate is $\approx 60\%$ at large separations
    ($>1~\pc$), but there are still a significant number of comoving pairs at
    these very wide separations.
    \emph{Right}: Distribution of chord lengths (tangential separations) for all
    comoving star pairs (gray) and only those identified as comoving with 3D
    velocity information (black).
    Noise in the parallax measurements makes the 3D separation distribution
    appear roughly uniform.
    Even the distribution in chord length appears to be approximately linear for
    large separations ($\gtrsim 0.1~\pc$).
    \label{fig:separation}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram of confirmed comoving star pairs plotted over
    the density of all \tgas\ sources within 200 pc; similar to
    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
    Several interesting pairs that include main-sequence and giant component
    stars are highlighted in the text
    (\sectionname~\ref{sec:interesting-pairs} and
    \figurename~\ref{fig:highlighted-cmd}).
    \label{fig:genuine-cmd}}
\end{figure}

An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have radial
velocities (for each star) reported in \DR{5} of the Radial Velocity Experiment
(RAVE; \citealt{Kunder:2017}).
Using the same procedure described in \sectionname~\ref{sec:catalog}, we
identify confirmed comoving pairs using the RAVE velocities for all pairs within
$200~\pc$ and find a similar false-positive rate at large separations ($\approx
38\%$ are confirmed).
Of the \nraveoverlap\ RAVE-overlap candidate comoving pairs (within $200~\pc$)
that meet the above criteria, we find \nravecomoving\ confirmed comoving pairs.
We plot the confirmed comoving pairs identified through radial velocity
follow-up in this work along with the RAVE-confirmed pairs in
\figurename~\ref{fig:separation-RAVE}.
Top histogram shows the separation (chord length) distribution for the joint
sample, scatter plot shows log-chord-length and the pair distances, and right
histogram shows the distance distribution of the confirmed pairs.
At larger distances ($\gtrsim 150~\pc$) it appears that the detection efficiency
of our method drops, as expected given the hard cut in log-FML-ratio applied to
define the candidate sample.
Reassuringly, there do not appear to be any trends of the chord length
distribution with distance.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
  \end{center}
  \caption{%
    \emph{Top}: Distribution of chord lengths for all confirmed comoving star
      pairs, either from the radial velocity follow-up presented in this work or
      from RAVE \DR{5} radial velocities.
      At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log)
      appears uniform, consistent with past measurements of the wide binary
      separation distribution.
      At separations between $\approx 10^{-1}$--$10^{0}~\pc$, there appears to
      be an underabundance of comoving pairs, but the number of comoving pairs
      then begins to increase again out to our sample separation limit of
      $10~\pc$.
    \emph{Middle}: Chord length plotted against mean distance for each comoving
      pair identified from this work (circles, blue) or from RAVE overlap
      (squares, orange).
    \emph{Right}: Distance distribution for all confirmed comoving pairs.
      Note that at distances $\gtrsim 150~\pc$, our detection efficiency appears
      to decline.
    \label{fig:separation-RAVE}}
\end{figure}

\subsection{A few highlighted comoving pairs}\label{sec:interesting-pairs}

@@

We compile Tycho 2 ($B$, $V$; \citealt{derp}), 2MASS ($J$, $H$, $K_{\rm s}$;
\citealt{derp}), and WISE ($W_1$, $W_2$, $W_3$; \citealt{derp}) photometry for
each star in the pairs below and fit isochrones to the stars individually to
measure their stellar parameters.
We use the \package{isochrones} \python\ package (\citealt{Morton:2015}) to fit
the photometric and parallax information (and uncertainties) for each star to
infer the \logg, \teff, and stellar mass; these values are reported in
\tablename~\ref{tbl:highlighted-pairs}.
In brief, we use the model grids from the Dartmouth Stellar Evolution Database
(\citealt{Dotter:2008}) and the built-in interpolation in \package{isochrones}
to compute the posterior probability of a given set of stellar parameters given
the data (photometry and parallax).
We generate samples (in stellar parameters) from this posterior pdf using the
ensemble MCMC sampler implemented in \package{emcee}
(\citealt{Foreman-Mackey:2013}) using 128 walkers.
We run the sampler for 256 initial steps from random initial positions, compute
the median position of the walkers in parameter-space and re-initialize the
walkers in a small Gaussian ball around this median position to run for another
1024 burn-in steps.
We then discard all but the last position and re-run for a final 8192 steps,
saving only every 128th step.
The stellar parameter values listed below are computed by taking the median
posterior sample, and the uncertainties are estimated by computing the median
absolute deviation, ${\rm MAD}$, and converting to a standard deviation under
the assumption of Gaussianity ($\sigma \approx 1.5 \times {\rm MAD}$).

\begin{table}[ht]
  \begin{center}
    \begin{tabular}{ c | c | c | c | c }
      \toprule
        Star name & \logg\ & \teff\ [K] & mass [$M_\odot$] &
          chord length [pc]  \\
        \toprule
        HD 95088 & $3.7 \pm 0.02$ & $5303 \pm 74$ & $1.490 \pm 0.002$ &
          \multirow{2}{*}{$8.6 \pm 0.8$}\\
        HD 92278 & $4.2 \pm 0.05$ & $8519 \pm 300$ & $1.8 \pm 0.1$ &\\
        \midrule
        HD 79247 & $3.61 \pm 0.02$ & $5075 \pm 60$ & $1.150 \pm 0.004$ &
          \multirow{2}{*}{$6.8 \pm 0.3$}\\
        TYC 3807-1439-1 & $4.50 \pm 0.04$ & $5409 \pm 80$ & $0.88 \pm 0.05$ &\\
        \midrule
        HD 54318 & $4.15 \pm 0.04$ & $10030 \pm 345$ & $2.4 \pm 0.1$ &
          \multirow{2}{*}{$8.4 \pm 1.1$}\\
        TYC 3792-658-1 & $4.54 \pm 0.03$ & $5462 \pm 139$ & $0.89 \pm 0.06$ &\\
      \bottomrule
    \end{tabular}
    \caption{
      Stellar parameters inferred using isochrone-fitting for the three pairs
      highlighted in \sectionname~\ref{sec:interesting-pairs}.
      \label{tbl:highlighted-pairs}
    }
  \end{center}
\end{table}

@@

From their surface gravities and temperatures, the component stars of this pair
appear to be an A-type dwarf (HD 92278) and a slightly less massive G-type
subgiant (HD 95088).
This pair is separated by $4.5^\circ$ on the sky (corresponding to $\approx
8.6~\pc$ in chord length, see \tablename~\ref{tbl:highlighted-pairs}) and have
consistent distances within their uncertainties of $\approx 110 \pm 5~\pc$.
At first glance, it seems unlikely that the two stars could be coeval, as the
more massive A star has not begun its post-main-sequence evolution.
One possible explanation is that this pair is a false-positive in three
dimensions, in which case they are unassociated but just appear to be comoving
with small relative velocity at present day.

Other speculative explanations for this pair stem from the possibility that it
is the outcome of an interaction between even more stars.
For example, an interaction between two wide binaries could result in an
exchange of members, a hardening of one resulting binary, and a widening of the
other binary (see, e.g., \citealt{Nathan??}).
Or, the initial system could have been an unstable three-body system that
resulted in a tight inner binary and a wide outer binary;
if the inner binary were close enough, it could have merged and formed a blue
straggler star (e.g., \citealt{??}), leaving the wide binary untouched.

High-resolution spectroscopy of these stars --- and precise measurements of the
stellar parameters and chemical abundances --- may help to distinguish these
scenarios.

@@

The components of this pair appear to be a K-type subgiant (HD 79247) and a
G-type dwarf (TYC 3807-1439-1).
This pair is separated by $4^\circ$ on the sky or $6.8~\pc$ in chord length and
have an apparent radial separation of $\approx 3~\pc$, corresponding to a 3D
separation of $\approx 9.8~\pc$ at a distance of $\approx 100~\pc$.
Coeval stars at different phases of stellar evolution are extremely valuable for
calibrating and testing stellar evolution models (e.g., \citealt{Torres:2013}).
With a larger sample of confirmed comoving pairs, identified pairs that contain
both an evolved star and a main sequence star --- such as this pair --- will be
of great interest.

@@

The components of this pair appear to be an A- or B-type dwarf (HD 54318) and a
G-type dwarf (TYC 3792-658-1), the largest magnitude difference between all of
the confirmed comoving pairs.
This pair is separated by $2.5^\circ$ on the sky or $8.5~\pc$ in chord length
and have a consistent distance of $\approx 190~\pc$ within their uncertainties.
The mass ratio distribution of comoving star pairs will help constrain star
formation models (e.g., \citealt{Raghavan:2010}).

@@

@@

% Low-resolution spectroscopic follow-up of comoving pairs.

% Using relative RV's, confirm a sample of very wide separation comoving star pairs.

% Either wide binaries, unbound binaries, or dissolving clusters

% Need much larger sample with no cuts on separation or signal-to-noise to study
% true separation distribution

@@

The information contained in a large sample of widely-separated, comoving star
pairs spans from the dynamics of stars and dark matter in the Galaxy, the
fundamentals of star formation, and the stability of planetary systems.
Through our own low-resolution spectroscopic follow-up and a cross-match with
the RAVE \DR{5} catalog, we have identified \ncomovingtotal\ comoving star pairs
by combining these radial velocity measurements with astrometry from the \tgas\
catalog.
The separation distribution of these spectroscopically-confirmed comoving star
pairs presented shows an abundance of comoving pairs at separations between
$1$--$10~\pc$ (\figurename~\ref{fig:separation-RAVE}), the imposed limit in our
initial search (\citealt{Oh:2017}).
This result is further tentative evidence (see also \citealt{Shaya:2011}) for
the existence of a population of unbound comoving stars that represent the
ultimate fate of wide binaries and moving groups that slowly dissociate due to
weak perturbations from stars (\citealt{Jiang:2010}) and other massive
perturbers in the Galaxy.
With future data releases from the \gaia\ mission combined with large
spectroscopic surveys, the anticipated catalogs of confirmed comoving star pairs
will offer a new ``stereo'' view on the evolution and state of our Galaxy.

@@

It is a pleasure to thank
Ruth Angus (Columbia),
John Brewer (Yale),
Jason Curtis (Columbia),
Keith Hawkins (Columbia),
David W. Hogg (NYU),
Melissa Ness (MPIA), and
Dave Zurek (AMNH)
for useful discussions and advice.
We thank James Davenport (WWU) for releasing his open source spectroscopic
reduction code, \package{PyDIS}, which served as inspiration for the pipeline
used in this work.
This work is based on observations obtained at the MDM Observatory, operated by
Dartmouth College, Columbia University, Ohio State University, Ohio University,
and the University of Michigan.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.
This research has made use of the SIMBAD database, operated at CDS, Strasbourg,
France.
% The authors are pleased to acknowledge that the work reported on in this
% paper was substantially performed at the TIGRESS high performance computer
% center at Princeton University which is jointly supported by the Princeton
% Institute for Computational Science and Engineering and the Princeton
% University Office of Information Technology's Research Computing department.
